Delocalization of Eigenvectors in Random Matrices

### 2025-03-20 Draft.tex
quantum diffusion

\subsection*{The Model}

The matrix model $(H_{ij})_{i, j\in \Gamma}$ that will be subject to our examination throughout this paper is defined as follows: identify $\Z_n$ with $\Gamma=\{1, \cdots, N\}$ and equip it with the aforementioned periodic distance $|i-j|_N$, s.t. $\forall_{|i-j|_N>W} H_{ij}=0$. Now, for a symmetric and compactly supported probability density function (PDF) on $\mathbb{R}$, define the doubly stochastic matrix $(S_{xy})_{x, y\in \Gamma}$ as: $$\mathbb{E}|H_{xy}|^2=S_{xy} := Z_{N, W}^{-1}f\left(\frac{|x-y|_N}{W}\right),$$
where $Z_{N, W}\asymp W$ is a normalizing constant bounded deterministically above and below by $W$. A technical, but necessary assumption is for $S$ to admit a matrix square root $S^{1/2}$, satisfying the same properties for a similarly symmetric and compactly supported PDF $f'$. \vspace{0.75 cm}
\begin{center}
\begin{tikzpicture}
    % Define parameters
    \def\n{4} % Matrix dimension
    \def\w{0.5} % Half-width of the band

@@

\subsection*{Useful identities}
Let us establish several facts and lemmas that we will need throughout our later work. We can begin by observing that: $$m(z)(m(z)+z) = \frac{1}{4}(-z+\sqrt{z^2-4})(z+\sqrt{z^2-4})=\frac{1}{4}(-z^2+z^2-4)=-1$$
$$\Rightarrow |m(z)| = \frac{1}{|m(z)+z|^{-1}}$$
Now, take the imaginary part of $(2)$: $$\text{Im}(m(z)) + \text{Im}\left(\frac{1}{m(z)}\right)+\eta=0$$
We can calculate the second term by simply expanding: $$\frac{1}{m(z)} = \frac{1}{\text{Re}(m(z))+\text{Im}(m(z))i} = \frac{\text{Re}(m(z))-\text{Im}(m(z))i}{|m(z)|^2}\Leftrightarrow \text{Im}\left(\frac{1}{m(z)}\right) = -\frac{\text{Im}(m(z))}{|m(z)|^2}$$
$$\Rightarrow\text{Im}(m(z)) + \text{Im}\left(\frac{1}{m(z)}\right)+\eta=\text{Im}(m(z)) -\frac{\text{Im}(m(z))}{|m(z)|^2}+\eta=0$$
$$\Rightarrow \text{Im}(m(z))\left(1-\frac{1}{|m(z)|^2}\right)=-\eta$$
And since $\text{Im}(m(z))>0$ and $\eta>0$, we must have:
\begin{equation*}\label{3.1} 1-\frac{1}{|m(z)|^2}<0\quad \Leftrightarrow\quad |m(z)|<1, \tag{3.1}
\end{equation*}
which is also equivalent to $\text{Im }m \asymp 1$, where $\asymp$ means being bounded above and below up to a fixed, positive factor. Furthermore, let us rewrite the latter equation as the following two identities: \begin{equation*}\label{3.2}1-|m(z)|^2 = \frac{\eta|m(z)|^2}{\text{Im}(m(z))}
\tag{3.2}
\end{equation*}
\begin{equation*}\label{3.2}\text{Im}(m(z))= \frac{\eta|m(z)|^2}{1-|m(z)|^2 }
\tag{3.3}
\end{equation*}
 \textcolor{red}{Write \hyperref[lemma-a1]{Lemma A.1} here or transfer}. Now, let $E\in [-10, 10]$ and $\eta\in (0, 10]$. First observe that $|m(z)|\geq c>0$ can never be zero, since it implies: $$m(z) = \frac{-z+\sqrt{z^2-4}}{2}=0 \Leftrightarrow \sqrt{z^2-4}=z$$
$$\Rightarrow z^2-4=z^2\Leftrightarrow -4=0$$

@@

$$\Theta := \frac{|m(z)|^2S}{1-|m(z)|^2S}$$
For $z$ in the bulk ($|E|<2$), by \hyperref[3.1]{(3.1)} and  \hyperref[lemma-a1]{Lemma A.1}, $1-|m(z)|^2\asymp \eta\Leftrightarrow \exists \alpha, 1-|m(z)|^2\leq \alpha\eta$, we have that: 
 $$\Theta =\frac{|m(z)|^2S}{1-|m(z)|^2S} = \frac{|m(z)|^2S}{1-|m(z)|^2+|m(z)|^2(S-\text{Id})}\leq   \frac{S}{1-|m(z)|^2+|m(z)|^2(S-\text{Id})}\sim \frac{S}{\alpha\eta+(S-\text{Id})}$$

>>>>

\subsection*{The Model}
\label{model}
The matrix model $(H_{ij})_{i, j\in \Gamma}$ that will be subject to our examination throughout this paper is defined as follows: identify $\Z_n$ with $\Gamma=\{1, \cdots, N\}$ and equip it with the aforementioned periodic distance $|i-j|_N$, s.t. $\forall_{|i-j|_N>W} H_{ij}=0$. Now, for a symmetric and compactly supported probability density function (PDF) on $\mathbb{R}$, define the doubly stochastic matrix $(S_{xy})_{x, y\in \Gamma}$ as: $$\mathbb{E}|H_{xy}|^2=S_{xy} := Z_{N, W}^{-1}f\left(\frac{|x-y|_N}{W}\right),$$
where $Z_{N, W}\asymp W$ is a normalizing constant bounded deterministically above and below by $W$. A technical, but necessary assumption is for $S$ to admit a matrix square root $S^{1/2}$, satisfying the same properties for a similarly symmetric and compactly supported PDF $f'$. \vspace{0.75 cm}
\begin{center}
\begin{tikzpicture}
    % Define parameters
    \def\n{4} % Matrix dimension
    \def\w{0.5} % Half-width of the band

@@

\subsection*{Useful identities}
Let us establish several facts and lemmas that we will need throughout our later work. We can begin by observing that: $$m(z)(m(z)+z) = \frac{1}{4}(-z+\sqrt{z^2-4})(z+\sqrt{z^2-4})=\frac{1}{4}(-z^2+z^2-4)=-1$$
$$\Rightarrow |m(z)| = \frac{1}{|m(z)+z|^{-1}}$$
Now, take the imaginary part of $(2)$: $$\text{Im}(m(z)) + \text{Im}\left(\frac{1}{m(z)}\right)+\eta=0$$
We can calculate the second term by simply expanding: $$\frac{1}{m(z)} = \frac{1}{\text{Re}(m(z))+\text{Im}(m(z))i} = \frac{\text{Re}(m(z))-\text{Im}(m(z))i}{|m(z)|^2}\Leftrightarrow \text{Im}\left(\frac{1}{m(z)}\right) = -\frac{\text{Im}(m(z))}{|m(z)|^2}$$
$$\Rightarrow\text{Im}(m(z)) + \text{Im}\left(\frac{1}{m(z)}\right)+\eta=\text{Im}(m(z)) -\frac{\text{Im}(m(z))}{|m(z)|^2}+\eta=0$$
$$\Rightarrow \text{Im}(m(z))\left(1-\frac{1}{|m(z)|^2}\right)=-\eta$$
And since $\text{Im}(m(z))>0$ and $\eta>0$, we must have:
\begin{equation*}\label{3.1} 1-\frac{1}{|m(z)|^2}<0\quad \Leftrightarrow\quad |m(z)|<1, \tag{3.1}
\end{equation*}
which is also equivalent to $\text{Im }m \asymp 1$, where $\asymp$ means being bounded above and below up to a fixed, positive factor. Furthermore, let us rewrite the latter equation as the following two identities: \begin{equation*}\label{3.2}1-|m(z)|^2 = \frac{\eta|m(z)|^2}{\text{Im}(m(z))}
\tag{3.2}
\end{equation*}
\begin{equation*}\label{3.2}\text{Im}(m(z))= \frac{\eta|m(z)|^2}{1-|m(z)|^2 }
\tag{3.3}
\end{equation*}
Now, let $E\in [-10, 10]$ and $\eta\in (0, 10]$. First observe that $|m(z)|\geq c>0$ can never be zero, since it implies: $$m(z) = \frac{-z+\sqrt{z^2-4}}{2}=0 \Leftrightarrow \sqrt{z^2-4}=z$$
$$\Rightarrow z^2-4=z^2\Leftrightarrow -4=0$$

@@

$$\Theta := \frac{|m(z)|^2S}{1-|m(z)|^2S}$$
The first theorem established by Dubova and Yang (Theorem 2 \cite{bandSDE}) shows that in the bulk $T\approx \Theta$:
\begin{boxtheorem}{Theorem 1: Quantum Diffusion}
For $|E|<2$ fixed, assume $\exists \nu>0$, s.t. $\eta \asymp W^2 N^{-2}$ and $W\geq W^{8/11+\eta}$. Then: $$\max_{x, y}|T_{xy}-\Theta_{xy}|\prec W^{-\frac{7}{4}}\eta^{-\frac{3}{2}}$$
\end{boxtheorem}
This result has a physical interpretation. First, observe
that by \hyperref[3.1]{(3.1)} and  \hyperref[lemma-a1]{Lemma A.1}, $1-|m(z)|^2\asymp \eta\Leftrightarrow \exists \alpha, 1-|m(z)|^2\leq \alpha\eta$, we have that: 
 $$\Theta =\frac{|m(z)|^2S}{1-|m(z)|^2S} = \frac{|m(z)|^2S}{1-|m(z)|^2-|m(z)|^2(S-\text{Id})}\sim \frac{S}{\alpha\eta-(S-\text{Id})}$$
\textcolor{red}{Shouldn't this be a minus sign? Theorem 2 in \cite{bandSDE} states it with a plus sign.}
If we recall the phenomenology of our \hyperref[model]{model}
